---
title: "synDataGenerator"
author: "Eliza Duvall"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r, include = FALSE}
library("corrplot")
library("igraph")
```

Functions to generate synthetic data - in my case, the synthetic data is a correlation matrix.  

## Function

Parameters in the function:  

- Lmin/Lmax = are the LOWEST correlation thresholds at a minimum and maximum (default is Lmin = 0.05 and Lmax = 0.3) 
- Mmin/Mmax = are the MID correlation thresholds at a minimum and maximum (default is Mmin = 0.4 and Mmax = 0.7) 
- Hmin/Hmax = are the HIGHEST correlation thresholds at a minimum and maximum (default is Hmin = 0.8 and Lmax = 0.99) 
- nodes = the number of nodes to include
- Lfraction = the fraction of the nodes that should fall in the LOWEST correlation category
- Mfraction = the fraction of the nodes that should fall in the MID correlation category
- Hfraction = the fraction of the nodes that should fall in the HIGHEST correlation category  

**NOTE:** The fractions should sum to 1

```{r}
synDataGenerator <- function(Lmin = 0.05, Lmax = 0.3,
                             Mmin = 0.4, Mmax = 0.7,
                             Hmin = 0.8, Hmax = 0.99,
                             nodes, Lfraction = 0.5, 
                             Mfraction = 0, Hfraction = 0.5){
  numVals = ((nodes*nodes)-nodes)/2 # the number of edges/correlation values that need to be calculated
  
  vector <- runif(numVals*Lfraction, Lmin, Lmax) # randomly choose values for lower
  vector <- append(vector, runif(numVals*Mfraction, Mmin, Mmax)) # for mid
  vector <- append(vector, runif(numVals*Hfraction, Hmin, Hmax)) # for high 
  
  mat <- matrix(0, nodes, nodes) # put together in a matrix
  mat[lower.tri(mat, diag=FALSE)] <- vector
  mat[upper.tri(mat, diag=FALSE)] <- vector
  
  return(mat)
}
```


## Test function default

Test this function

```{r, warning=FALSE}
defaultMatrix <- synDataGenerator(nodes = 10)
defaultMatrix
```

And let's look at this correlation where color represents correlation strength

```{r}
corrplot(defaultMatrix, method = "number")
```

Generate a correlation matrix where there are blocks of highly correlation nodes

```{r}
m <- synDataGenerator(node = 22, Lfraction = 1, Hfraction = 0)
matH1 <- synDataGenerator(nodes = 10, Lfraction = 0, Hfraction = 1)
matH2 <- synDataGenerator(nodes = 5, Lfraction = 0, Hfraction = 1)
matH3 <- synDataGenerator(nodes = 7, Lfraction = 0, Hfraction = 1)

m[1:10, 1:10] <- matH1
m[11:15, 11:15] <- matH2
m[16:22, 16:22] <- matH3
```

## Generate highly correlated blocks

Let's look at the correlation matrix in a nice plot, and we will see blocks of highly correlated notes

```{r}
corrplot(m, method = "number")
```

The purpose of creating this synthetic correlation is for network analysis. Let's use this correlation matrix as an adjacency matrix and see what the complete network looks like. **Note,** a complete network is one in which all nodes are connected to all nodes.  

```{r, warning=FALSE}
g2 <- graph_from_adjacency_matrix(m, mode = "undirected", weighted = TRUE)
E(g2)$width <- E(g2)$weight*3
plot(g2)
```

And here we see the three neighborhoods/blocks of highly correlated values :) 



